---
title: "Rworksheet_Gregorio#4b"
output: pdf_document
date: "2023-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r}
vectorA <- c(1,2,3,4,5)

matrixA <- matrix(0,nrow = 5, ncol =5)

for (i in 1:5)
  for (j in 1:5)
  {
    matrixA[i,j] <- abs (vectorA[i] - vectorA[j])
  }

matrixA

```
#2.
```{r, echo = TRUE}

for (i in 1:5) {
  cat(paste0("\"", rep("*", i), "\""), "\n")
}
```
#3.
n <- as.numeric(readline(prompt = "Enter a number to start the Fibonacci sequence: "))

a <- 0
b <- 1
c <- a + b

repeat {
  if (c > 500) { 
    break
  }
  if (a == 0 & b == 1) {
    cat(b, " ")
  }
  cat(c, " ")
  a <- b
  b <- c
  c <- a + b
}

4.Import the dataset as shown in Figure 1 you have created previously.
4a.What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result
```{r}
ShoesData <- read.csv("Shoe sizes.csv")
head(ShoesData)
```
4b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female?  Write the R scripts and its output.

```{r}

maleSub <- subset(ShoesData, Gender == "M")
femSub <- subset(ShoesData, Gender == "F")

cat("The number of observation in male subset:", nrow(maleSub),"\n")
cat("The number of observation in female subset:", nrow(femSub),"\n")

```

4c. Create a graph for the number of males and females for Household Data. Use plot(), chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and its result

```{r}

GenderMF <- table(ShoesData$Gender)

barplot(GenderMF,
        main = "Number of Male and Female in Household Data",
        xlab = "Gender",
        ylab = "Count" ,
        col = c("black", "green"),
        legend.text = c("Female", "Male"))

```








#5.
```{r}
spending_data <- data.frame(
  Category = c("Food", "Electricity", "Savings", "Miscellaneous"),
  Value = c(60, 10, 5, 25)
)

spending_data$Percentage <- spending_data$Value / sum(spending_data$Value) * 100

colors <- c("black", "green", "red", "violet")

pie(spending_data$Value,
    labels = paste(spending_data$Category,"(",spending_data$Percentage,"%)"),
    col = colors,
    main = "Monthly Income Spending of Dela Cruz Family")

legend("topright", spending_data$Category, fill = colors)
```

6. Use the iris dataset
```{r}
data(iris)
data
```

6a. Check for the structure of the dataset using the str() function. Describe what you have seen in the output.
```{r}

str(iris)
# The dataset contains information on iris blossoms. It contains information on the length and width of the sepal as well as the length and width of the petal for 150 different iris flowers. It also keeps track of the iris species to which every flower belongs. The dataset contains three species: setosa, versicolor, and virginica.
```

6b.Create an R object that will contain the mean of the sepal.length, sepal.width,petal.length,and petal.width. What is the R script and its result?

```{r}
meanOfFlowers <- colMeans(iris[,1:4])
meanOfFlowers
```

6c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.
```{r}
species_count <- table(iris$Species)
pie(species_count, labels = species_count, col = rainbow(length(species_count)), main = "Species Distribution")
legend("topright", names(species_count), cex = 0.8, fill = rainbow(length(species_count)))

```

6d.Subset the species into setosa, versicolor, and virginica. Write the R scripts and show the last six (6) rows of each species.
```{r}
# Subset the iris data set into the three species.
setosa_subset <- subset(iris, Species == "setosa")
versicolor_subset <- subset(iris, Species == "versicolor")
virginica_subset <- subset(iris, Species == "virginica")

setosa_subset
versicolor_subset
virginica_subset

# Display the last six rows of each species.
tail(setosa_subset, 6)
tail(versicolor_subset, 6)
tail(virginica_subset, 6)
```

6e. Create a scatterplot of the sepal.length and sepal.width using the different species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.

```{r}
# Convert the "Species" column to a factor
iris$Species <- as.factor(iris$Species)

# Create a scatterplot
plot(
  Sepal.Length ~ Sepal.Width,
  data = iris,
  pch = as.integer(iris$Species),  # Use different pch symbols for each species
  col = as.integer(iris$Species),  # Use different colors for each species
  xlab = "Sepal Length",
  ylab = "Sepal Width",
  main = "Iris Dataset",
  sub = "Sepal width and length"
)

# Add a legend
legend("topright", legend = levels(iris$Species), col = 1:3, pch = 1:3, cex = 0.8, title = "Species")
```

6f.Interpret the result.
```{r}
# The dataset consists of five variables (columns) and 150 observations (rows) in a data frame format.
# Petal.Length, Petal.Width, Sepal.Length, and Sepal. Width are the names of the four numerical variables. These stand for the measures of the length and width of the petals, respectively.
# The factor variable Species, which represents the species of iris flowers, is the sixth variable. There are three tiers to it: "setosa," "versicolor," and "virginica," denoting the many iris flower species included in the dataset.
```
7.Import the alexa-file.xlsx. Check on the variations. Notice that there are extra whitespaces among black variants (Black Dot, Black Plus, Black Show, Black Spot). Also on the white variants (White Dot, White Plus, White Show, White Spot).

```{r}

library(readxl)
alexa_file <- read_excel("alexa_file.xlsx")
alexa_file

```

7a. Rename the white and black variants by using gsub() function.
```{r}
alexa_file$variation <- gsub("Black  Dot", "BlackDot", alexa_file$variation)
alexa_file$variation <- gsub("Black  Plus", "BlackPlus", alexa_file$variation)
alexa_file$variation <- gsub("Black  Show", "BlackShow", alexa_file$variation)
alexa_file$variation <- gsub("Black  Spot", "BlackSpot", alexa_file$variation)

alexa_file$variation <- gsub("White  Dot", "WhiteDot", alexa_file$variation)
alexa_file$variation <- gsub("White  Plus", "WhitePlus", alexa_file$variation)
alexa_file$variation <- gsub("White  Show", "WhiteShow", alexa_file$variation)
alexa_file$variation <- gsub("White  Spot", "WhiteSpot", alexa_file$variation)

alexa_file
```

7b.  Get the total number of each variations and save it into another object. Save the object as variations.RData. Write the R scripts. What is its result?
```{r}
library("dplyr")

var_total <- alexa_file %>%
  count(alexa_file$variation)

var_total
save(var_total, file = "var_total.RData")

```

7c. From the variations.RData, create a barplot(). Complete the details of the chart which include the title, color, labels of each bar.
```{r}
load("var_total.RData")
var_total
varNames <- var_total$`alexa_file$variation`

totalPlot <- barplot(var_total$n,
        names.arg = varNames,
        main = "Total number of each variations",
        xlab = "Name of Variations",
        ylab = "Total of each Variatons",
        col = 1:16,
        space = 2)

```

```

7d. Create a barplot() for the black and white variations. Plot it in 1 frame, side by side. Complete the details of the chart.

```{r}

blackVars <- var_total[var_total$`alexa_file$variation` %in% c("Black", "BlackPlus" , "BlackShow" ,"BlackSpot" , "BlackDot"), ]

whiteVars <- var_total[var_total$`alexa_file$variation` %in% c("White", "WhiteDot", "WhitePlus", "WhiteShow", " WhiteSpot"), ]

par(mfrow = c(1,2))

barplot(height = blackVars$n,
        names.arg = blackVars$`alexa_file$variation`,
        col = c("black"),
        main = "Black Variations",
        xlab = "Variation",
        ylab = "Count",
        border = "black")

barplot(height = whiteVars$n,
        names.arg = whiteVars$`alexa_file$variation`,
        col = c("black"),
        main = "White Variations",
        xlab = "Variation",
        ylab = "Count",
        border = "black")
```


