---
title: "Rworksheet_Gregorio#4c"
author: "Ardwayne B. Gregorio  2 - C"
date: "2023-11-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg

1a. Show your solutions on how to import a csv file into the environment.
```{r}

mpgdata <- read.csv("mpg.csv")

```


1b. Which variables from mpg dataset are categorical?
```{r}
str(mpgdata)




```

c.Which are continuos variables?
```{r}
#displ: Engine displacement in liters.
#hwy: Highway miles per gallon.
#cty: City miles per gallon.
```

2.
a.
```{r}
library(ggplot2)


manumodel <- table(mpg$manufacturer)

manumax <- names(which.max(manumodel))

cat("Manufacturer with the most models:", manumax, "\n")


#finding most variations
manumaxdata <- subset(mpg, manufacturer==manumax)
model_counts <- table(manumaxdata$model)

mostvarmod <- names(which.max(model_counts))
cat("Model with the most variations:", mostvarmod, "\n")

```
2a.Group the manufacturers and find the unique models. Show your codes and result.
```{r}
library(ggplot2)
library(dplyr)

manumod <- mpg%>%
  group_by(manufacturer)%>%
  summarize(uniquemodel = n_distinct(model))

print(manumod)



```
2b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}

#ggplot
library(ggplot2)
ggplot(manumod, aes(x=manufacturer,y = uniquemodel))+
  geom_bar(stat = "Identity",fill = "blue")+
  labs(title = "Unique model per manufacturer",
       x = "Manufacturer",
       y = "Number of unique models")+
  theme_gray()

   


```

2a.What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
library(ggplot2)
ggplot(mpg, aes(model, manufacturer)) + geom_point()

# it shows the points where model and manufacturer are.

```
2b. For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}
#Indeed, that is helpful.

```
3.
```{r}
library(ggplot2)

top20 <- head(mpg,20)

ggplot(top20, aes(x = model, y = year))+
  geom_point()+
  labs(title = "Model vs Year(Top 20 observations)",
       x = "Model",
       y= "Year")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))




```
4.Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}
library(dplyr)
modelgroup <- mpg%>%
  group_by(manufacturer)
 

print(modelgroup)

```

4a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and
colors. Show code and results.
```{r}
library(dplyr)
library(ggplot2)

data(mpgdata)

top20 <- head(mpg,20)

ggplot(top20, aes(x = model, fill = model)) +
  geom_bar() +
  labs(title = "Number of Cars per Model (Top 20 Observations)",
       x = "Model",
       y = "Number of Cars") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  scale_fill_manual(values = rainbow(nrow(top20))) 


```
4b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}
library(dplyr)
library(ggplot2)

data(mpgdata)

top20 <- head(mpg,20)

ggplot(top20, aes(x = model, fill = model)) +
  geom_bar() +
  labs(title = "Number of Cars per Model (Top 20 Observations)",
       x = "Number of Cars",
       y = "Model") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = rainbow(nrow(top20))) +
  coord_flip()



```

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.
```{r}
library(ggplot2)

data(mpgdata)


ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  scale_color_continuous(name = "Engine Displacement") +
  theme_minimal()

```
5a. How would you describe its relationship? Show the codes and its result.
```{r}
relation_cyldispl <- cor(mpg$cyl , mpg$displ)

cat("Correlation Coefficient:", relation_cyldispl, "\n")


library(ggplot2)
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  scale_color_continuous(name = "Engine Displacement") +
  theme_minimal()

```
6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
output?
 
```{r}

ggplot(mpg, aes(x = displ, y = hwy, color = year)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement",
       y = "Highway MPG") +
  scale_color_gradient(low="green",high="red") +
  theme_minimal()


```
6. Import the traffic.csv onto your R environment.

6a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.

```{r}
trafficdata <- read.csv("traffic.csv")

number_obs <- nrow(trafficdata)
var_traffic <- ncol(trafficdata)
varnames <- names(trafficdata)

cat("Number of Observations:", number_obs, "\n")
cat("Number of Variables:", var_traffic, "\n")
cat("Variables:", paste(varnames, collapse = ", "), "\n")

```
6b. subset the traffic dataset into junctions. What is the R codes and its output?
```{r}
junction_traffic <- c(1,2,3,4)


junction_subset <- trafficdata[trafficdata$Junction %in% junction_traffic, ]

junction_subset


```


6c. Plot each junction in a using geom_line(). Show your solution and output.

```{r}
library(ggplot2)

ggplot(trafficdata, aes(x = DateTime, y = Vehicles, color = as.factor(Junction))) +
  geom_line() +
  scale_color_manual(values = c("red", "orange", "green", "blue")) + 
  labs(title = "Traffic by Junction",
       x = "Date Time",
       y = "Number of Vehicles") +
  theme_minimal() +
  facet_wrap(~Junction, scales = "free_y", ncol = 1) +
  guides(color = guide_legend(title = "Junction"))



```
7. From alexa_file.xlsx, import it to your environment
```{r}
library(readxl)

alexafile <- read_excel("alexa_file.xlsx")
```

7a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.

```{r}
alexa_obser <- nrow(alexafile)

alexacol_obser <- ncol(alexafile)



cat("The number of observations on alexa is:",alexa_obser,"\n")
cat("The number of coloumns on alexa is:",alexacol_obser,"\n")

```


7b.group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.
```{r}
library(dplyr)
groupvari <- alexafile%>%
  group_by(variation)%>%
  summarise(totalcount_ = n())
 

groupvari


```
7c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.
```{r}
library(ggplot2)

ggplot(alexafile, aes(x = variation)) +
  geom_bar() +
  labs(title = "Distribution of Variations",
       x = "Variation",
       y = "Count") +
  theme_minimal()
```

7d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.
```{r}
alexafile$date <- as.Date(alexafile$date)

alexafile$month <- format(alexafile$date, "%m")

monthscount <- alexafile %>%
  group_by(month) %>%
  summarise(num_reviews = n())


monthlyreview <- table(monthscount)

ggplot(monthscount, aes(x = month, y = num_reviews, group = 1)) +
  geom_line(color = "blue") +
  labs(title = "Number of Verified Reviews Per Month",
       x = "Month of 2017", y = "Number of Reviews")

```

7e. Get the relationship of variations and ratings. Which variations got the most h ghest in rating? Plot a
graph to show its relationship. Show your solution and answer.
```{r}
library(dplyr)
ggplot(alexafile, aes(x = variation, y = rating, fill = variation)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  scale_fill_manual(values = rainbow(n = length(unique(alexafile$variation)))) +  
  labs(title = "Relationship Between Variations and Ratings",
       x = "Variations",
       y = "Average Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

var_ratings <-  alexafile %>%
  group_by(variation)%>%
  summarise(avearage_rating = mean(rating, na.rm = TRUE))

max_rating <- max(var_ratings$average_rating, na.rm = TRUE)

highrate <- alexafile %>%
  filter(rating == max_rating)
print(highrate)

```

